

First of all, I am deeply grateful to Andrea Ferrara, who has been my supervisor and mentor for the last three years. His help, teaching, and advice have been invaluable to me. I first met him when I did not know anything about research, and since then, he has never stopped conducting me on the path to becoming a scientist. A huge thank you also to my co-supervisor, Andrea Pallottini. He helped me a lot with the project, and I could always rely on him to assist me and answer my numerous questions. In addition, my sincere gratitude goes to my internal supervisor, Michele Cignoni, for his precious feedback on my work and insightful comments on the very first draft of this document. 

\vspace{0.5cm}
This thesis work develops from a project that has started more than two years ago. In this project, I was helped and advised by several people: Simona Gallerani, Livia Vallini, Davide Decataldo, and Seiji Fujimoto. In particular, I want to thank Davide for helping me implement the cooling function, and Seiji, for sharing his study of the ALPINE \CII emission profiles. 


\vspace{0.5cm}
Being part of the Cosmology Group at SNS, even just for a few months, was one of the best parts of this project. This is because of the people I had the pleasure to meet there. I want to mention my office mates, Lorenzo and Daniele, and the other Ph.D. students here in the group, especially Ivan, David, Laura, and Zip. Thanks for all the meals and the coffee breaks we spent together.


\vspace{0.5cm}
To me, this thesis work is not just the ending of a project. It is also the ending of a five-year experience in Pisa. For this reason, I want to take this opportunity to say thank you to the people I have shared this journey with: my class at SNS, and all the friends I have met at SNS and UniPi. They are too many to be listed here, but it's only because of them if I feel utterly grateful for the time spent here in Pisa. A special thanks to Beppe, for reading this thesis; to my flatmates Marco, Massimo, Federico, Lavinia, and Margherita, for the time shared, especially during the lockdown; and to Seyma, for accepting my poor judgment in hiking trips, and for giving me a reason to graduate on time.


\vspace{0.5cm}
My final thanks go to my family: my parents, siblings, grandparents, aunts, uncles, and cousins. Most of all, I want to thank my parents, Enrica and Carlo. Their advice, constant support, and unconditional love have been essential to me even in these years spent far from home. 
